
This analysis relies on all the codes which filenames start by "\texttt{Bayes}" from the \textbf{/Scripts-R/} folder of the github repository. Al the functions created that are used and are also made available through the package are in \textbf{/R/\texttt{BayesFunc.R}}.


"It is often the case that more than one model provides an adequate fit to the data. Sensitivity analysis determines by what extent posterior inferences change when alternative
models are used"   book risk analysis other section pp.2.

"The basic method of sensitivity analysis is to fit several models to
the same problem. Posterior inferences from each model can then be compared."


\section{From \texttt{evdbayes} R package : MH algorithm}

The \texttt{evdbayes} is the only package available on CRAN for (see \citet{ribatet_users_2006}) is a very old package which has not been updated since a while. We had problems to understand both its structure and the famous "black-box"


\section{From Our Functions (R package)}



\section{From HMC algorithm using STAN language}

The problem is maybe from \ref{likgevintro}. The parameter $\xi$ is relatively near the region that could be problematic, causing convergence issues. 


\section{Comparisons}

\citet{hartmann_bayesian_2016}
We can calculate the effective sample size (ESS) using the posterior samples for each parameter : 

\begin{equation}
\text{ESS}=N\cdot (1+2\sum_k\gamma(k))^{-1}
\end{equation}

where $N$ is still the number of posterior samples and $\gamma(k)$ are the monotone lag $k$ sample autocorrelations. \citet{geyer__1992}. We can thus interpret this as the number of effectively independent samples. 


\subsection{STAN}

Benefits : 

\begin{itemize}
	\item Allows more flexibility (?) through the mathematical formulation of the formula
	\item It is really smoother and clearer (straightforward) for this kind of problems 
\end{itemize}

Drawbacks : 

\begin{itemize}
	\item New language with all the problems/errors arising when learning it. 
\end{itemize}



\section{Comparison with frequentists results}